Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-06 paper/paper.tex
edited the cplx-overview section, moved a paragraph (to be edited) into appendix

In this section we overview $\mathbb{C}$-valued networks, their motivation and peculiarities.
In out exposition we shall focus mostly on linear layers, keeping in mind that transferring 
the ideas to convolutional layers is straightforward.

The most recent study of complex-valued networks \cite{trabelsi_deep_2017}, provides a set
of building blocks for deep $\cplx$-valued networks. The core of their work is structured
framework for complex-valued operations including representation and linear layers (dense
and convolutional) as well as $\cplx$-valued activations and complex batch-normalization
and weight initialization.

Following \cite{trabelsi_deep_2017}, we naturally identify $\cplx$ with $\real^2$ via $
  z_\mathrm{r} + j z_\mathrm{i} \mapsto (z_\mathrm{r}, z_\mathrm{i})
$, storing the real and imaginary parts in interleaved format, and emulate $\cplx$-algebra
using $\real$-valued arithmetic. Thus designed $\cplx$-network is, essentially, a
$\real$-network with \textit{double} the layers' widths to accommodate real and imaginary
parts and with constraints on layers' parameters that make them respect $\cplx$-algebra.
This makes it possible to retrofit $\cplx$-valued networks into the existing $\real$-valued
auto-differentiation libraries.

In particular, common linear layers, such as dense layers and convolutions,
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
$$
\real^{2 \times \mathrm{[in]}}
  \to \real^{2 \times \mathrm{[out]}}
  \colon (x_r, x_i)
    \mapsto \Bigl(
      L_r x_r - L_i x_i,
      L_r x_i + L_i x_r
    \Bigr)
  \,, $$
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are the unique $\real$-linear operators that make up $L = L_r + j L_i$ the overall
effect of the layer. For instance, to respect $\cplx$-arithmetic a convolutional layer with
$\cplx$-kernel $W$ can be implemented using two $\real$-convolutions with $\real$-kernels
$\Re{W}$ and $\Im{W}$. Other algebraic operations, such as division and roots in batch
normalization \cite{trabelsi_deep_2017}, are implemented in such a way as to respect
$\cplx$-algebra.
%
Concerning activations, \cite{trabelsi_deep_2017} consider the following: independent
componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$; $\cplx$-differentiable exponential and trigonometric activations; and \textit{modRelu}
activation, which preserves the angle and applies soft-thresholding to the modulus: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - \theta)_+
  % z \mapsto z \bigl(
  %     1 - \tfrac\theta{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable parameter $\theta\in \real$.

The question of differentiability of complex-valued networks is tricky, since they usually
use non-differentiable non-linearities. However, a more severe problem stems from the loss
objective being real-valued. Indeed, if $f\colon \cplx \to \cplx$ were $\cplx$-differentiable
(holomorphic), the composition of $f$ with a non-trivial $h \colon \cplx\to \real$ makes it
not $\cplx$-differentiable with respect parameters, since $h$ violates Cauchy-Riemann conditions.
This problem is dealt with by using Wirtinger, or $\cplx\real$ calculus, which subsumes
$\cplx$-calculus, and is applicable to non-holomorphic functions of $\cplx$-argument,
\cite{adali_complex-valued_2011}. As a basis for backpropagation algorithm it was considered
in \cite{benvenuto_complex_1992} and \cite{trabelsi_deep_2017}.

Basically $\cplx\real$ calculus regard $f(z)$ as a function of two independent variables $z$ 
and its conjugate $\conj{z}$ and symbolically defines derivative operators $
  \tfrac{\partial}{\partial z}
$ and $
  \tfrac{\partial}{\partial \conj{z}}
$, that satisfy linearity, product and chain rules, and respect complex conjugation. The important
property is that, if we identify $f(z) = f(u + jv) = F(u, v)$, then the $\cplx\real$ calculus
rules differential of $f$ as a $\cplx$ function coincides its differential as an $\real^2$
function. This makes it possible to use and apply the rules of $\real$-calculus to the networks
represented in paired $\real$ fashion and use existing $\real$-valued backpropagation in autograd
software. The key analytic advantage is that computing derivatives is more streamlined, less
messy, and derivative expressions involving complex variables turn out to be simpler that their
real-imaginary analogues, \cite{adali_complex-valued_2011}.

@@

% TLDR: we use Wirtinger calculus!

>>>>

In this section we overview $\mathbb{C}$-valued networks, their motivation and peculiarities.
In our exposition we shall focus mostly on linear layers, keeping in mind that transferring 
the ideas to convolutional layers is straightforward.

The study \cite{trabelsi_deep_2017}, outlines building blocks for deep $\cplx$-valued
networks and describes suitable representation and operations including linear layers (dense
and convolutional), $\cplx$-valued activations, complex batch-normalization and suitable
weight initialization. In fact all papers touching one way or another upon the subject of
$\cplx$-valued networks, refrain each other, and the current study is no exception.

Following \cite{trabelsi_deep_2017}, we naturally identify $\cplx$ with $\real^2$ via $
  z_\mathrm{r} + j z_\mathrm{i} \mapsto (z_\mathrm{r}, z_\mathrm{i})
$, storing the real and imaginary parts in interleaved format, and emulate $\cplx$-algebra
using $\real$-valued arithmetic. In particular, convolutional and linear layers,
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation} \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique linear operators that make up $L = L_r + j L_i$. Thus, a $\cplx$-convolutional
layer with complex kernel $W$ can be implemented as two $\real$-convolutions with kernels
$\Re{W}$ and $\Im{W}$ using \eqref{eq:cplx-lin-op}. Other algebraic operations, such as
trigonometric functions, and complex division and roots in batch normalization
\cite{trabelsi_deep_2017}, are implemented through their counterparts in $\real^2$. This
makes $\cplx$-valued network into an intricately connected $\real$-valued computational
graph that respects $\cplx$-algebra.

Concerning activations, \cite{trabelsi_deep_2017} consider the following: independent
componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$; $\cplx$-differentiable exponential and trigonometric activations; and \textit{modRelu}
activation, which preserves the angle and applies soft-thresholding to the modulus: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - \theta)_+
  % z \mapsto z \bigl(
  %     1 - \tfrac\theta{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable parameter $\theta\in \real$.

Complex networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed, as implied by Cauchy-Riemann conditions, a
$\cplx$-differentiable real-valued function is necessarily trivial. This issue is dealt with
by employing Wirtinger, or $\cplx\real$ calculus, which generalizes holomorphic calculus
to non-holomorphic functions of $\cplx$-argument, \cite{adali_complex-valued_2011}. It was
considered as a basis for $\cplx$ version of backpropagation in \cite{benvenuto_complex_1992}
and \cite{trabelsi_deep_2017} (see sec.~\ref{sub:wirtinger_calculus} for details). The key
benefit of $\cplx\real$ calculus is straightforward symbolic differentiation, and the fact,
that this allows retrofiting $\cplx$-valued networks into existing $\real$ deep learning
auto-differentiation libraries.

\todo{read on holomorphic nets}

@@

% TLDR: we use Wirtinger calculus!
Basically $\cplx\real$ calculus regard $f(z)$ as a function of two independent variables $z$ 
and its conjugate $\conj{z}$ and symbolically defines derivative operators $
  \tfrac{\partial}{\partial z}
$ and $
  \tfrac{\partial}{\partial \conj{z}}
$, that satisfy linearity, product and chain rules, and respect complex conjugation. The important
property is that, if we identify $f(z) = f(u + jv) = F(u, v)$, then the $\cplx\real$ calculus
rules differential of $f$ as a $\cplx$ function coincides its differential as an $\real^2$
function. This makes it possible to use and apply the rules of $\real$-calculus to the networks
represented in paired $\real$ fashion and use existing $\real$-valued backpropagation in autograd
software. The key analytic advantage is that computing derivatives is more streamlined, less
messy, and derivative expressions involving complex variables turn out to be simpler that their
real-imaginary analogues, \cite{adali_complex-valued_2011}.
